---
title: "Interpreting Data Visualizations"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css"]
    lib_dir: libs
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"

---

class: title-slide

```{r echo = FALSE, message=FALSE}
library(openintro)
library(tidyverse)
theme_set(theme_gray(base_size = 18))
```

<br>
<br>
.right-panel[ 
<br>

# `r rmarkdown::metadata$title`

## `r rmarkdown::metadata$author`

]

---

class: middle

## Reminder

- Close all apps on your computer other than zoom.  
- Open slides for this session from the cluster website (https://uci-dshs.netlify.app/). 


---

class: middle

[Limbo Lines: Dead Here, Alive There](https://pudding.cool/2018/02/death/)

[Anxiety: It Gets Worse, Then It Gets Better](https://www.instagram.com/p/B4U81DJlqTG/)

[The Wealth & Health of Nations](https://observablehq.com/@mbostock/the-wealth-health-of-nations)

---

class: middle

Data Visualizations

- are graphical representations of data

--

- use different colors, shapes, and the coordinate system to summarize data

--

- tell a story

--

- are useful for exploring data







---

class:inverse middle

.font75[Visuals with a Single Categorical Variable]

---

## Barplots

---

class:inverse middle

.font75[Visuals with a Single Numeric Variable]

---

## Boxplots

---

## Histograms


---

class: middle

[Exploring Histograms Interactively](http://tinlizzie.org/histograms/)

---

class: middle center

[There is no "best" number of bins](https://en.wikipedia.org/wiki/Histogram#Number_of_bins_and_width)

---

class: inverse middle center

.font75[Visuals with Two Categorical Variables]

---

## Standardized Bar Plot

```{r echo=FALSE, fig.align='center', fig.height=4}
ggplot(data = babies,
       aes(x = smoke, 
           fill = parity)) + 
  geom_bar(position = "fill") +
  ylab("Proportion") +
  xlab("Mother's Smoker Status")
```

---

## Dodged Bar Plot

```{r echo=FALSE, fig.align='center', fig.height=4}
ggplot(data = babies,
       aes(x = smoke, 
           fill = parity)) + 
  geom_bar(position = "dodge")
```



---

class: middle inverse

.font75[Visuals with a single numerical and single categorical variable.]

---

class: inverse middle

.font75[Visuals with Two Numerical Variables]

---


## Barplots

Simple
Dodged

---

## Boxplots


---

## Histogram
Binwidth

Right skewed,
left-skewed
Do ordering of mean median

---

## Scatterplots
