---
title: "Changing Variables"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css"]
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide



```{r echo = FALSE, message = FALSE}
library(tidyverse)
options(scipen = 999)
```


<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

```{r echo = FALSE, message = FALSE}
library(tidyverse)

arthritis <- read_csv("https://raw.githubusercontent.com/cosmos-uci-dshs/data/main/RheumArth_Tx_AgeComparisons.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = case_when(sex == 0 ~ "female",
                   sex == 1 ~ "male")) %>% 
  mutate(sex = as.factor(sex)) %>% 
  mutate(age_gp = case_when(age_gp == 1 ~ "control",
                   age_gp == 2 ~ "elderly")) %>% 
  mutate(age_gp = as.factor(age_gp)) %>% 
  mutate(cdai_yn = case_when(cdai_yn == 1 ~ "no",
                             cdai_yn == 2 ~ "yes")) %>%
  mutate(cdai_yn = as.factor(cdai_yn))


arthritis <- 
  arthritis %>% 
  select(age, age_gp, 
         sex, yrs_from_dx,
         cdai)
```



```{r}
glimpse(arthritis)
```

**Goal**: 

Create a new variable called `age_months` that represents `age` in months.

---

```{r}
arthritis %>% 
  mutate(age_months = age*12)
```

---

```{r}
glimpse(arthritis)
```

---

**Goal**: 

Create a new variable called `cdai_level` which reflects the levels in the the data dictionary

- Remission ≤ 2.8
- Low Disease Activity > 2.8 and ≤ 10
- Moderate Disease Activity > 10 and ≤ 22
- High Disease Activity > 22

---

```{r}
arthritis %>% 
  mutate(cdai_level = case_when(
    cdai <=2.8 ~ "Remission", 
    cdai > 2.8 & cdai <= 10 ~ "Low",
    cdai > 10 & cdai <=22  ~ "Moderate",
    cdai > 22 ~ "High")) 
```

---

We can't really see what we have created 




---

We can use pipes with ggplot too!

.left-panel[
```{r eval = FALSE}
arthritis %>% 
  mutate(cdai_level = case_when(
    cdai <=2.8 ~ "Remission", 
    cdai > 2.8 & cdai <= 10 ~ "Low",
    cdai > 10 & cdai <=22  ~ "Moderate",
    cdai > 22 ~ "High")) %>% 
  ggplot(aes(x = cdai_level)) +
  geom_bar()
```
]

.right-panel[
```{r echo = FALSE}
arthritis %>% 
  mutate(cdai_level = case_when(
    cdai <=2.8 ~ "Remission", 
    cdai > 2.8 & cdai <= 10 ~ "Low",
    cdai > 10 & cdai <=22  ~ "Moderate",
    cdai > 22 ~ "High")) %>% 
  ggplot(aes(x = cdai_level)) +
  geom_bar()
```
]

---


```{r}
arthritis <- arthritis %>% 
  mutate(cdai_level = case_when(
    cdai <=2.8 ~ "Remission", 
    cdai > 2.8 & cdai <= 10 ~ "Low",
    cdai > 10 & cdai <=22  ~ "Moderate",
    cdai > 22 ~ "High")) 
```

---

```{r}
glimpse(arthritis)
```

**Goal**: 

Make `cdai_level` a factor and `age` an integer. 

---

```{r}
arthritis %>% 
  mutate(cdai_level = as.factor(cdai_level),
         age = as.integer(age)) 
```

---

class: middle

`as.factor()` - makes a vector factor  
`as.numeric()` - makes a vector numeric  
`as.integer()` - makes a vector integer  
`as.double()` - makes a vector double  
`as.character()` - makes a vector character  










